Perihelion advance of Mercury

### 2019-06-17 perihelion_advance_of_mercury.Rmd
theoretical analysis

We need to isolate the change points (actually calculating aphelion rather than perihelion for convencence):

@@

And to convert to astronomical units we use
```{r}
conversion <-
  c(
      seconds_per_minute = 60,
      minutes_per_hour   = 60,
      hours_per_day      = 24,
      days_per_year      = 365.25,
      years_per_century  = 100,
      degrees_per_radian = 180/pi,
      minutes_per_degree = 60,
      seconds_per_minute = 60
  )

```

@@

>>>>

We need to isolate the change points (actually calculating aphelion
rather than perihelion for convenience):

@@

And to convert to astronomical units we use
```{r}
conversion <-
  c(
      seconds_per_minute = 60,
      minutes_per_hour   = 60,
      hours_per_day      = 24,
      days_per_year      = 365.25,
      years_per_century  = 100,
      degrees_per_radian = 180/pi,
      minutes_per_degree = 60,
      seconds_per_minute = 60
  )
```

@@

##  Theoretical analysis

In current notation, a first-order analysis gives

\[
\frac{24\pi^3a^2}{t^2c^2\left(1-e^2\right)}
\]

for the precession.  In R idiom:

```{r}
precession <- 24*pi^3*a^2/(per^2*sol^2*(1-ecc^2))  # radians/orbit
precession
```

To convert this figure which is given in radians per orbit) to
arcseconds per century, we need to multiply by a sequence of factors:

```{r}
seconds_per_orbit <- prod(c(
    seconds_per_minute = 60,
    minutes_per_hour   = 60,
    hours_per_day      = 24,
    days_per_orbit     = 87.969
))

seconds_per_century <- prod(c(
    seconds_per_minute = 60,
    minutes_per_hour   = 60,
    hours_per_day      = 24,
    days_per_year      = 365.25,
    years_per_century  = 100
))

orbits_per_century <- seconds_per_century/seconds_per_orbit

arcseconds_per_radian <- prod(
    c(
        degrees_per_radian = 180/pi,
        minutes_per_degree = 60,
        seconds_per_minute = 60
   	     ))

arcseconds_per_orbit <- precession*arcseconds_per_radian
arcseconds_per_century <- arcseconds_per_orbit * orbits_per_century
arcseconds_per_century
```

Compare with the value of about 43.37 obtained from numerical
integration, showing close agreement.
